---
title: "the final project"
subtitle: "Project report"
author: "abdullah sami"
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float: yes
    highlight: zenburn
    df_print: paged
  word_document:
    toc: yes
  pdf_document:
    toc: yes
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo= TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# 1. Introduction to Rmarkdown

Rmarkdown file (.Rmd) allows to:

-   Save and execute code

-   Generate high quality reports that can be shared with an audience, by knitting together plots, tables, and results with narrative text, and rendering to a variety of formats like HTML, PDF, MS Word, or MS PowerPoint.

# 2. How to make a report using Rmarkdown?

1- Create new file: File \> New File \> R Markdown

2- Create code chunks: Run code by line, by chunk, or all at once.

3- Write text and add tables, figures, images

4- Customize header: set output format

5- Save and knit your document

6- Publish (rpubs.com, Rstudio connect etc)

# 3. Data analysis

## 3.1 Dataset description

We want to understand the factors on which the pricing of cars depends. Specifically, we want to know how well those variables describe the price of a car.

To do so, we got a dataset from [Kaggle](https://www.kaggle.com/datasets/hellbuoy/car-price-prediction?resource=download).

The dataset has 205 observations and 26 columns.

```{r}
# libraries
library(tidyverse)
library(knitr)
```

```{r}
# import data from csv file
cars.price <- read.csv("archive/CarPrice.csv")

# examine data structure
str(cars.price)

# head of the data
head(cars.price, 5)
tail(cars.price, 5)
```

We select some variables to use in our analysis.

```{r}
# selecting columns
# method 1
cars.price_2 <- cars.price[,3:26]

# method 2: eliminate columns 1 and 2
cars.price_3 <- cars.price[,-c(1,2)]

# method 3: select variables by name
cars.price_4 <- cars.price[,c("stroke", "price","carbody")]

```

```{r}
# check fuel type values
table(cars.price$fueltype) %>% kable(col.names = c("Value","Frequency"), caption="Fuel type values")

# method 4: select variables & filter rows (dplyr)
cars.price_5 <- cars.price %>%
  select(stroke, price, carbody, peakrpm, CarName, fueltype, drivewheel, curbweight, enginetype, horsepower) %>%
  filter(fueltype == "gas")
```

```{r}
# summary statistics numeric variables
summary(cars.price_5[,c(1,2,4,8,10)]) %>% kable(caption = "Summary statistics numeric variables")
```

```{r}
# convert character variables to factors for one variable:
cars.price_5$carbody <- as.factor(cars.price_5$carbody)

# convert 5 variables from character to factor
cars.price_5[,-c(1,2,4,8,10)] <- lapply(cars.price_5[,-c(1,2,4,8,10)] , as.factor)

# summary of factor variables
options(knitr.kable.NA = '')
summary(cars.price_5[,-c(1,2,4,8,10)]) %>% kable(caption="Summary of factor variables")
```

## 3.2. Data visualization

### 3.2.1. Univariate plots

```{r fig.align='center'}

# use ggplot to make a histogram
ggplot(data = cars.price_5, aes(x=price)) + # select data and x variable
  geom_histogram(binwidth = 2000, color="white", fill="orange") +
  # change binwidth, bars border and fill color
  xlab("Price ($)") + # add label to the x axis
  ylab("Count") + # add label to the y axis
  ggtitle("Histogram of car prices") + # add title to the plot
  theme_minimal() # add a theme

```

```{r fig.align='center'}

# use ggplot to make a histogram
ggplot(data = cars.price_5, aes(y=stroke)) + # select data and x variable
  geom_boxplot(fill="orange") + # fill color
  ylab("Stroke") + # add label to the x axis
  #ylab("Count") + # add label to the y axis
  ggtitle("Boxplot of stroke") + # add title to the plot
  theme_minimal() # add a theme

```
